---
title: "LC QC Report"
output:
  html_document: default
  pdf_document: default
date version: May 31, 2019
---



```{r Global Options, echo=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.align='center', warning=FALSE, message=FALSE, error=TRUE)
options(tinytex.verbose = TRUE)
```


## Total ion chromatogram

```{r Extract TICs, eval=TRUE, echo=FALSE}

cat('Query TIC total area ', signif(100*sum(query_TIC_load)/sum(reference_TIC_load), digits=4), '% of reference.',
    sep='')

cat('Est. amount injected ', signif(AMOUNT_INJECTED*sum(query_TIC_load)/sum(reference_TIC_load), digits=4), ' ng.',
    sep='')

plot(reference_TIC_load[,1], reference_TIC_load[,2], type="h", lwd=1, col='green', 
     xlab='retention time (min)', ylab='total ion current')
lines(query_TIC_load[,1], query_TIC_load[,2], type="h", lwd=1, col='blue')

```
<center>Figure 1: Total ion chromatograms of reference (green) and query (blue). </center>

```{r System contaminants, eval=TRUE, echo=FALSE}
# under development
cat('Evaluation of system contaminants such as polysiloxane and detergents, is under development.', sep='')

```

***

## Peak characteristics

Here we estimate peak width using lag-1 autocorrelation.
```{r Estimate peakWidth, eval=TRUE, echo=FALSE}
cat('Mean peak width (FWHM equivalent) =', signif(peak_width, digits=4), "seconds")

```
The mean peak width is calculated but is not used as a quality metric. 


```{r Estimate peakCapacity, eval=TRUE, echo=FALSE}
cat('Theor. peak capacity =', signif(gradient_length/peak_width, digits=4))

```
The peak capacity should be greater than 300 (for most nano systems), while between 200 and 300 is marginal, and less than 200 is considered bad.

***

## Retention Coefficients

Here we train a linear model (Palmblad *et al*., 2002) for predicting retention times, and compare the amino acid coefficients of the query model with those from the reference model. This will reveal changes in mobile phase composition, such as ionic strength or pH. The deviation of these coeffecients from the diagonal implies your mobile phase pH may be off. The correlation between query and reference coeffecients is used as a diagnostic feature. Also, if the line has a change in y-intercept then this may indicate your pumps aren't working. 

```{r Compare retentionCoefficients, eval=TRUE, echo=FALSE}
plot(reference_coeffs[1:21], query_coeffs[1:21], pch=19, cex=2.5, 
     col=alpha(aa_colors, 0.6), xlab='reference retention coefficient (Z-score)', 
     ylab='query retention coefficient (Z-score)', xlim=limits, ylim=limits, text(reference_coeffs[1:21], query_coeffs[1:21]+0.0047*(max(query_coeffs[1:21])-min(query_coeffs[1:21])), aa_olc, col='black', font=2, cex=0.8)) 

abline(a=0, b=1, lty=2)

```
<center>Figure 2: Retention time coefficients for query compared with reference. The closer to the 45-degree diagonal the better. Deviation in basic and/or acidic residues suggests different pH or ionic strength.</center>