---
title: "LC QC"
date version: "April 24, 2019"
output:
  html_document: default
---

This sub-report contains metrics related to the performance of LC system components. Metrics include estimated pH of mobile phases, column heater,  shape of gradient, peak shape statistics (such as FWHM for 25, 50 and 75% quartiles and base peak width). Future work will include determining % TIC that is polysiloxane or detergents in order to evaluate column health and system contamination.




```{r Global Options, include=TRUE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.align='center', warning=FALSE, message=FALSE, error=TRUE)
options(tinytex.verbose = TRUE)

```


#### 1. Load R packages and specify the search parameters and sequence database

First we load R packages, specify the parameters and sequence database:

```{r Setup}
library(mzR)
library(scales)

# Commented out lines must be set before render call
#QUERY <- '2018-10-18_HeLa_OTOT'                             # query for QC
#OUTPUT_DIR <- '/tmp/lkjdflkj'
#MZML_DIR

# The following are hard coded for now
REFERENCE <- '/data/2018-10-15_HeLa_OTOT'                    # reference (could be local)
AMOUNT_INJECTED <- 200                                       # ng HeLa injected (reference)
SPECTRAST_LIBRARY <- '/data/HeLa.splib'                      # HeLa library
FASTA_FILE <- '/data/up000005640.fasta'                      # sequence database
mzml_file <- paste(MZML_DIR,'/',QUERY,'.mzML',sep='')
cat('mzML file is',mzml_file)


```

### 3. Look up instrument manufacturer and model from query mzML file

Here we read in instrument metadata from the mzML file using the [mzR](http://bioconductor.org/packages/release/bioc/html/mzR.html) package. 

```{r Extract instrumentInfo}
query_file <- openMSfile(mzml_file)
make <- instrumentInfo(query_file)$manufacturer
model <- instrumentInfo(query_file)$model
cat('Query file is', make, model)   # look up make and model of file for QC    
```