---
title: "Sample QC"
date version: "April 24, 2019"
output:
  html_document: default
---

This sub-report contains metrics related to QC sample integrity. The QC Benchmarker automatically detects whether the sample was HeLa (Thermo), yeast (Promega) or *E. coli* (Waters) digest. Metrics include estimated sample loading, % spectra matched to sample, and +1 charge states over the gradient. Future work will include determining matches to cRAP/PRIDE contaminants.




```{r Global Options, include=TRUE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.align='center', warning=FALSE, message=FALSE, error=TRUE)
options(tinytex.verbose = TRUE)

```


#### 1. Load R packages and specify the search parameters and sequence database

First we load R packages, specify the parameters and sequence database:

```{r Setup}
library(mzR)
library(scales)

# Commented out lines must be set before render call
#QUERY <- '2018-10-18_HeLa_OTOT'                             # query for QC
#OUTPUT_DIR <- '/tmp/lkjdflkj'
#MZML_DIR

# The following are hard coded for now
REFERENCE <- '/data/2018-10-15_HeLa_OTOT'                    # reference (could be local)
AMOUNT_INJECTED <- 200                                       # ng HeLa injected (reference)
SPECTRAST_LIBRARY <- '/data/HeLa.splib'                      # HeLa library
FASTA_FILE <- '/data/up000005640.fasta'                      # sequence database
mzml_file <- paste(MZML_DIR,'/',QUERY,'.mzML',sep='')
cat('mzML file is',mzml_file)


```

### 3. Look up instrument manufacturer and model from query mzML file

Here we read in instrument metadata from the mzML file using the [mzR](http://bioconductor.org/packages/release/bioc/html/mzR.html) package. 

```{r Extract instrumentInfo}
query_file <- openMSfile(mzml_file)
make <- instrumentInfo(query_file)$manufacturer
model <- instrumentInfo(query_file)$model
cat('Query file is', make, model)   # look up make and model of file for QC    
```